A Survey of Possible Exoskeleton Control Architectures and Algorithms\\ (Draft 3.0)

### 2015-09-25 control.tex
cga updates

\subsection{Full Atlas robot implementation.}

Discuss or point to DRC papers by multiple groups.

>>>>

\subsection{Full robot implementations.}

Point to DRC papers by multiple groups.

Figure~\ref{f:cga1} shows the architecture of our robot control based on hierarchical optimization. 
It was successfully tested on our SARCOS Primus Humanoid
and ATLAS (Boston Dynamics) platforms. Our ATLAS robot demonstrated rough terrain walking, ladder climbing, and full body manipulation in the DARPA Robotics Challenge (DRC) Trials of 2013. This work also received 
a ``Best Oral Paper Award'' at Humanoids 2013~\cite{Feng/etal13}. Figure~\ref{fig:terrain_pic} shows a sequence of snapshots of ATLAS 
walking on rough terrain and plots of measured trajectories of the feet and the center of mass (CoM) for the DRC Trials.
Figure~\ref{fig:ladder_pic} shows similar data for ladder climbing.
In the DARPA Robotics Challenge Finals in 2015 our robot was the only
biped that did not fall down or need to be rescued by 
humans during the two days of the evaluations~\cite{cga-hum15}.
Our robot performance was also the most consistent across the two days.
Because of the way the evaluations were run, we have only limited data
from the actual evaluations. Videos of our robot's performance during the
DRC Finals are available~\cite{cga-web-page}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[t!]
\centering
\includegraphics[width=0.75\textwidth]{Figures/cga1_modified}
\caption{Controller architecture based on hierarchical optimization.}
\label{f:cga1}
\end{figure}

\begin{figure}[t]
  \begin{center}
    {\includegraphics[height=3.8cm]{Figures/atlas_terrain}}
    \caption{ATLAS robot practicing for the terrain task of the DRC. Photo snapshots taken every 5 seconds. Measured trajectories in the transverse (a) and sagittal planes (b) of the left (red) and right foot (green) and the center of mass (CoM, blue) are shown. The robot walks in a straight line along $y=0$ in reality. Our state estimator at that time drifted significantly. This has been fixed by adding vision-based SLAM.
      }\label{fig:terrain_data} 
\label{fig:terrain_pic} 
  \end{center}
\end{figure}

\begin{figure}[t]
  \begin{center}
    {\includegraphics[height=4cm]{Figures/atlas_ladder}}
    \caption{
      ATLAS climbing the top half of the same 
      ladder as used in the DRC. Photo snapshots taken every 13 seconds. The
      top row shows repositioning of the hook hands, and the bottom row shows
      stepping up one tread.
      The two plots show ATLAS climbing the first five treads during
      the actual run at the DRC Trials. Axis definitions and color codes for foot and CoM trajectories are the same as in Fig.~\ref{fig:terrain_data}. Left and right hand positions are plotted with cyan and magenta dashed lines. 
      }\label{fig:ladder_data} 
\label{fig:ladder_pic} 
  \end{center}
\end{figure}

In the controller architecture, the {\bf Behavior Library} (Fig.~\ref{f:cga1}) contains
learned behaviors and behaviors created from demonstration or
motion capture of humans or other robots, and also acts as a cache
for planning. The output of the library is used to prime, speed up,
and guide the real time optimization process, and can be motion
or force trajectories in joint or task coordinates ($\vx_d(t)$), as well as
possibly including value, heuristic cost, or utility functions ($V_x(t)$)
that can be used
to guide search and optimization. 
Libraries can be built in advance, or when task specification are known.
The rest of the controller can
also operate without information from the library if suitable information
is not available.

The {\bf Multistep Low-Order Dynamics Planner} uses
receding horizon control to continually optimize trajectories of
motion and force of a simplified model of the current task.
In walking, this is usually center of mass ($COM_d(t)$),
angular momentum ($\mL_d(t)$),
contact force ($\mbox{contacts}_d(t)$) trajectories,
and value, heuristic cost, or utility functions ($V_c(t)$)
specifying tradeoffs between these quantities.
We optimize with a time horizon of several seconds
using Differential Dynamic Programming~\cite{ddp}.

For walking, our planner is similar in spirit to Kajita's Preview Control for a Linear Inverted
Pendulum Model (LIPM)~\cite{kajita03}. We are also using a CoM
model, reason about Zero Moment Point (ZMP), and use future information 
to guide the current trajectory. However, our planner can be generalized to 
nonlinear models and adjust foot steps while optimizing the CoM 
trajectory. 
%We explicitly add the vertical height in our  CoM model to handle height variations on rough terrain. 
Like capture point methods~\cite{capture_pt}, we take the 
next few steps into consideration but
do not plan to come to rest at the end. Ogura et al.~\cite{wabian2} 
investigated generating human-like walking with heel-strike and toe-off by 
parameterizing the swing foot trajectory. In contrast, we use very simple rules to guide the low level controller to achieve the same behaviors.

The {\bf Full Body QP Optimization} uses quadratic programming (QP) with a
very short time horizon (1ms on SARCOS and 2ms on ATLAS) to optimize the 
desired full state of the robot ($\vtheta_d$, $\dot{\vtheta}_d$),
joint torques ($\vtau_d$), and contact forces ($\vf_d$).
Currently, we pre-specify gain matrices 
($\mK_{\theta}$, $\mK_{\dot{\theta}}$, $\mK_{\tau}$,
$\mK_{\mbox{f}}$)
for the robot controller, but plan to explore how to optimize gains
as well as trajectories.
This module enforces constraints
such as robot kinematics and dynamics 
and joint, velocity, torque, actuator, friction cone,
and center of pressure limits, and resolves redundancies.
The module performs inverse kinematics and dynamics to provide us with
compliant motions and dynamic behaviors, while compensating for the
effects of modeling errors.

For the QP, we use a formulation developed in our group 
\cite{Stephens-thesis,Whitman10,Whitman-thesis}. Unlike~\cite{mistry10,usc_id1} who 
use orthogonal decomposition to project the allowable motions into the null 
space of the constraint Jacobian, and minimize costs in the contact constraints and the commands, we directly 
optimize a quadratic cost in terms of state accelerations, torques and contact 
forces on the full robot model. This design choice allows us to trade off physical quantities of 
interest. We are also able to directly reason about inequality constraints 
such as center of pressure within the support polygon, friction, and torque 
limits. Although it becomes a bigger QP problem, we are still able to solve it 
in real time. Hutter el al.~\cite{eth_id} resolved redundancy in inverse 
dynamics using a kinematic task prioritization approach that ensures lower 
priority tasks always exist in the null space of higher priority ones. In 
contrast to their strictly hierarchical approach, we minimize a sum of 
weighted terms. 
We can directly specify the relative importance of the terms by adjusting 
the weights.

%One popular approach to controlling humanoid robots is through inverse 
%kinematics with stiff joint position tracking. On the other hand, 
%inverse dynamics
%based approaches have gained increasing acceptance by providing compliant 
%motions and robustness to external perturbations. 
%However, the performance of inverse dynamics methods 
%is heavily dependent on high quality dynamic models, which are often very 
%difficult to produce for a physical robot. In contrast, inverse
%kinematics based approaches 
%only require kinematic models, which are much easier to generate in practice. 
%This was one reason we supplemented our original two part implementation with 
%an inverse kinematics controller for the physical robot.

The {\bf Low-Level Robot Controller} is a blend of our work and controllers
provided by the robot manufacturers. For SARCOS
and ATLAS, hydraulic valve commands for each
joint $i$ are generated by high rate servo controllers provided
by the manufacturers,
\begin{equation}
v_i = k_{\theta,i} ( \theta_i - \theta_{d,i} )
+ k_{\dot{\theta},i} ( \dot{\theta}_i - \dot{\theta}_{d,i} )
+ k_{\tau,i} ( \tau_i - \tau_{d,i} )
+ v_{\ff,i}
\end{equation}
where $v_i$ is the valve command,
$\theta_i$ is the joint angle,
$\dot{\theta}_i$ is the joint angular velocity,
$\tau_i$ is the joint torque,
and $v_{\ff,i}$ is a feedforward valve command.
Subscripts $d$ indicate desired values, and
quantities denoted with $k$ are scalar gains.
This joint level servo runs at 5kHz on SARCOS and at
1kHz on ATLAS.

We augment this independent joint control by coupling joints
and providing force control based on force/torque sensors in the wrists,
ankles, and skin of the robot,
\begin{equation}
\vv = \mK_{\theta} ( \vtheta - \vtheta_{d} )
+ \mK_{\dot{\theta}} ( \dot{\vtheta} - \dot{\vtheta}_{d} )
+ \mK_{\tau} ( \vtau - \vtau_{d} )
+ \mK_{\mbox{f}} ( \vf - \vf_{d} ) 
+ \vv_{\ff}
\end{equation}
where $\vv$, $\vtheta$, $\dot{\vtheta}$, $\vtau$, and $\vv_{\ff}$
are the corresponding vector quantities, and the $\mK$ are matrix gains
coupling all joints. $\vf$ are sensed contact forces.
The coupled controller runs at 1kHz on SARCOS and 500Hz on ATLAS.

Figure~\ref{f:cga1} does not include our state estimator, which provides
appropriate state information to all other modules. The details of this work on state estimation are
described in~\cite{xinji_com_est,Xinjilefu-thesis}.
